---
title: "Shifting Seas, Shifting Boundaries: Dynamic Marine Protected Area designs for a Changing Climate"
author: "Tim Cashion^1^ *, Tu Nguyen^2^, Talya ten Brink^3^ ^, Anne Mook^4^ ^, Juliano Palacios-Abrantes^5^ ^, Sarah M. Roberts^6^ ^"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  bookdown::html_document2
  #bookdown::word_document2
bibliography: MyCollection.bib
always_allow_html: yes
biblio-style: "apalike"
---

*Corresponding author. t.cashion@oceans.ubc.ca
^ These authors contributed equally. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning= F, message=F)
ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}
packages <- c(
  "tidyverse", #data manipulation and plotting
  "wesanderson", #Awesome colour schemes
  "rmarkdown", #rmarkdown document formatting and functions
  "knitr" #for printing tables with 'kable' 
)
ipak(packages)
options(knitr.kable.NA = '')

source("./R/end_of_century_average.R")
source("./R/spatial_maps_segment_plots.R")
```

```{r data, echo=FALSE, warning=FALSE, message=FALSE}


#For data cleaning files (converting EwE outputs to tidy format), see R script of 'OutputDataCleaning.R' 

mpa_labels <- tibble(design = c("MPA_Horiz_Static", 
                                "MPA_Square_Static", 
                                "MPA_Square_Shifting", 
                                "MPA_Vert_Static", 
                                "MPA_Network_Static", 
                                "MPA_Network_Shifting", 
                                "NoMPA"),
                     mpa_label = c("Horizontal Static", 
                                "Square Static", 
                                "Square Shifting", 
                                "Vertical Static", 
                                "Network Static", 
                                "Network Shifting", 
                                "No MPA"))

average_data <- data.table::fread("/nfs/mpasandclimatechange-data/Data/AnchovyBay/DataOutputs/AnnualAverageData_tidy.csv")
average_data <- average_data %>% 
  rename(year=Year) %>% 
  mutate(year=year-1) %>% 
  filter(scenario==10)
mpa_data <- read_csv("/nfs/mpasandclimatechange-data/Data/AnchovyBay/DataOutputs/MPAData_tidy.csv")
#temporarily limit to 5% percent scenario before building out others.
mpa_data <- mpa_data %>%
  mutate(scenario = gsub(scenario, pattern="MPA Scenario ", replacement = "")) %>% 
  mutate(scenario = gsub(scenario, pattern=" Percent", replacement = "")) %>% 
  mutate(scenario = as.numeric(scenario)) %>% 
  filter(scenario == 10)

# spatial_data_raw <- read_csv("nfs/mpasandclimatechange/data/AnchovyBay/DataOutputs/SpatialData_tidy.csv") #This file is large and takes time to load. Any way to truncate for quicker analysis and outputs? 
spatial_data <- data.table::fread("/nfs/mpasandclimatechange-data/Data/AnchovyBay/DataOutputs/SpatialData_tidy_10.csv")
price_data <- read_csv("/nfs/mpasandclimatechange-data/Data/AnchovyBay/DataOutputs/PriceData_tidy.csv")
#Need to revisit elasticity data as it seems very off
price_data$elasticity[price_data$elasticity>=2] <- 1.99

port_data <- read_csv("/nfs/mpasandclimatechange-data/Data/AnchovyBay/DataOutputs/PortData_tidy.csv")

average_fleet_data <- average_data %>% 
  filter(grepl(group, pattern="\\|"))

#Price elasticity:
orig_catch <- average_fleet_data %>% 
  filter(year==1 & RCP=="RCP2.6" & design=="NoMPA" & scenario=="10") %>% 
  select(group, value) %>% 
  rename(orig_catch = value)
average_fleet_data <- average_fleet_data %>% 
  left_join(price_data) %>% 
  left_join(orig_catch) %>% 
  rename(catch=value) %>% 
  mutate(PercChange = ((catch-orig_catch)/orig_catch))

average_fleet_data$PercChange[average_fleet_data$PercChange>=0.35] <- 0.35
average_fleet_data$PercChange[average_fleet_data$PercChange<=-0.35] <- -0.35


average_fleet_data <- average_fleet_data %>% 
  mutate(CurrentPrice = Price*(((-1*elasticity)*PercChange)+1))  %>% 
  mutate(value=CurrentPrice*catch)

elasticity_adjusted_prices <- average_fleet_data %>% 
  select(year, fg, design, RCP, CurrentPrice)

spatial_catch_data <- spatial_data %>% 
  filter(type=="Catch") %>% 
  left_join(elasticity_adjusted_prices) %>%
  rename(catch=value) %>% 
  mutate(value=CurrentPrice*catch)

#unique(spatial_data$type)
mpa_legend <- tibble(mpa=c(0,1,2), MPA=c("Outside", "Inside", "Adjacent"))
#Be careful with plotting the spatial data. To achieve the originally designed visual version in plots, the row_num (i.e. the y-axis) needs to be reversed as the original layout goes from 1 at the top to 20 at the bottom. Plots by default have higher values at the top and lower values at the bottom (i.e., the opposite of our grid layout). 

```

```{r mpa-spatial-prep}
mpa_spatial <- left_join(spatial_data, mpa_data) 
mpa_spatial$mpa[which(mpa_spatial$design=="NoMPA")] <- 0
test <- mpa_spatial[is.na(mpa_spatial$mpa),] 
mpa_spatial <- mpa_spatial %>% 
  left_join(mpa_legend)
mpa_spatial$MPA <- as.factor(mpa_spatial$MPA)
mpa_spatial$MPA <- fct_relevel(mpa_spatial$MPA, "Outside", after=0)
mpa_spatial$MPA <- fct_relevel(mpa_spatial$MPA, "Inside", after=1)

```

```{r plot-design}

zissou_continuous <- wes_palette(name = "Zissou1", n=100, type = c("continuous"))
zissou_discrete_3 <- wes_palette(name = "Zissou1")[c(1,3,5)]
#print(zissou_discrete_3)
theme_set(theme_classic())
zissou_blue <- wes_palette("Zissou1")[1]
zissou_red <- wes_palette("Zissou1")[5]


```


# Methods
Our aim was to explore the theoretical benefits of dynamic marine protected areas (MPAs) to respond to the effects of climate change. To undertake this analysis, we develop a spatially explicit model in Ecospace, the spatial form of Ecopath with Ecosim (EwE) [@walters_ecospace:_1999]. We adapted the theoretical ecosystem of 'Anchovy Bay' to test our analysis [@christensen_tutorial_2018]. This choice resulted from the finding that there are very few developed Ecospace models due to high data requirements. In addition, the theoretical aspect of this issue in future forecasting of climate change impacts lends itself well to a theoretical ecosystem where we can separate out the effects of climate change and MPA effects from the remaining uncertainty of ecosystem model parameterization. See supplementary materials for a full specification of the model. 

## Data
The functional groups were set to have environmental responses to climate change forcing functions. Dispersal values were used from a previous study that covered most of the functional groups in the ecosystem [@beattie_marine_2001]. The default price values from the original ecopath model were used [@christensen_tutorial_2018] (Table \@ref(tab:price)). Price elasticity was incorporated into the models based on known price elasticities of similar species/commercial groups. The values were sourced from a synthesis of same product elasticity [@asche_demand_2005] and an average was taken where multiple values were reported for a functional group (Table \@ref(tab:price-elasticity)). This was applied within the EwE model so that fishing effort behaviour would respond to changes in profitability of their fishing operations. We believe this allows a more realistic response of fishers to the revenue and profits of fishing when coupled with distance based fishing costs.   

```{r price}
price <- read_csv("./InputData/Sardine Bank-Off-vessel price.csv")
price$`Group name`[which(price$`Group name`=="Mackerel ad")] <- "Mackerel adult"
price <- price %>% 
  filter(`Group name` %in% c("Seals", "Cod", "Whiting", "Mackerel adult", "Anchovy", "Shrimp")) %>% 
  select(-c(X1, X8))
colnames(price) <- gsub(colnames(price), pattern="EUR", replacement= "$ ")
kable(price, caption="Ex-vessel prices used for this study. Obtained from Christensen, 2018.")
```

```{r price-elasticity}
price_sens <- read_csv("./InputData/pricesens.csv")
kable(price_sens, caption="Literature review of price elasticities corresponding to functional groups in this study. Obtained from Asche et al. 2005. Dependent Variable: p:price; q: quantity.")
```


## Scenarios  

### Climate change scenarios  
The model was constructed to compare the results of different MPA designs and protection strategies under climate change. We adopted two approximations for climate change effects on sea surface temperature based on the results of the Intergovernmental Panel on Climate Change (IPCC) estimates. These broadly encompass a 2$\circ$ and 4$\circ$ change in sea surface temperature by the end of the century (2100).      

### MPA designs
We then developed seven different MPA designs including a reference design without an MPA. We had four static MPAs to explore the spatial design of them and how this may react to climate change with a square MPA, a narrow vertical MPA, a narrow horizontal MPA, and a network of MPAs (Figure \@ref(fig:static-mpa-visuals)). Additionally, we had two dynamic MPAs that are the same dimensions as the square and network static MPAs hat shifted by one cell width (20 km) every 20 years (Figure \@ref(fig:shifting-mpa-visuals)). We differentiate between the narrow horizontal and narrow vertical MPAs as we hypothesize that vertically oriented MPAs will be more likely to benefit species as they shift poleward due to climate change. All designs with MPAs had the same amount of spatial area closed (e.g, 42 out of 400 cells, roughly 10% of the ecosystem). 

```{r static-mpa-visuals, fig.cap="Static MPA designs"}


mpa_visual <- mpa_data %>% 
  left_join(mpa_legend) %>% 
  left_join(mpa_labels) %>% 
  filter(!(grepl(design, pattern="Shifting"))) %>% 
  filter(design != "MPA_NoMPA") %>% 
  ggplot(aes(x=col_num, y=rev(row_num), fill=as.factor(MPA))) +
  geom_tile(color="grey") +
  facet_wrap(~mpa_label) +
  ylab("") + 
  xlab("") +
  scale_color_manual(values="grey") + 
  scale_fill_manual(values = rev(zissou_discrete_3)) +
  labs(fill="Area type") + 
  NULL
print(mpa_visual)
ggsave("./VisualOutputs/Static MPA Designs.png", dpi=300, width=8.5, height=11, units="in")
```

```{r shifting-mpa-visuals, fig.cap="Shifting MPA designs with caption indicating Year"}

mpa_visual_shifting <- mpa_data %>% 
  left_join(mpa_legend) %>% 
  left_join(mpa_labels) %>% 
  filter(grepl(design, pattern="Shifting")) %>% 
  filter(year %in% c(0, 20, 40, 60, 80)) %>% 
  ggplot(aes(x=col_num, y=rev(row_num), fill=as.factor(MPA))) +
  geom_tile(color="grey") +
  facet_wrap(~mpa_label+year, nrow=2) +
  ylab("") + 
  xlab("") +
  scale_color_manual(values="grey") + 
  scale_fill_manual(values = rev(zissou_discrete_3)) +
  labs(fill="Area type") + 
  NULL
print(mpa_visual_shifting)
ggsave("./VisualOutputs/Dynamic MPA Designs.png", dpi=300, width=8.5, height=11, units="in")
```

## Sensitivity analysis
Our sensitivity analysis is carried out by varying the base dispersal rate, which regulates the MPA spillover effect. To estimate the sensitivity to this parameter we varied this value by 25% and 50% in either direction (i.e., lower or higher base dispersal rates). We ran this second sensitivity analysis on three of our MPA designs (no MPAs, network static, and network shifting) under the 4$\circ$ warming scenario.

## Supplemental methods

### Model construction 

The EwE model used in this paper is an adaptation of an existing ecospace model. Ecospace models begin with the parameterization of a static mass-balanced model (Ecopath), modified by a temporal scenario (Ecosim), and finally spatialized to a certain habitat (Ecospace). 

#### Ecopath

The baseline model parameters and food web structure of Anchovy Bay Ecospace model were accepted [@christensen_tutorial_2018]. This model is made up of 11 functional groups, one of which is multi-stanza (has two age classes: juveniles and adults). The basic input parameters of the ecopath model can be found in Table \@ref(tab:ecopath) and the diet matrix in Table \@ref(tab:diets).  

```{r ecopath}
ecopath <- read_csv("./InputData/anchovybay_basicinput.csv", na="")
options(knitr.kable.NA = '')
kable(ecopath, caption= "Basic input parameters for Anchovy Bay ecopath model (Christensen, 2018).")
```


```{r diets}
diets <- read_csv("./InputData/anchovybay_dietmatrix.csv")
options(knitr.kable.NA = '')
colnames(diets)[1] <- "Group Number"
kable(diets, caption= "Diet matrix for Anchovy Bay ecopath model (Christensen, 2018).")
```

#### Ecosim
Forcing functions were applied for two climate change scenarios (2 degrees and 4 degrees warming). These were applied to consumers and producers and were assumed to affect their search rate. 

#### Ecospace
We defined environmental response parameters for species based on depth and temperature preferences. We then applied the forcing function of changes to sea surface temperature to the functional group specific response for the appropriate climate change scenario. To mimic a temperature gradient at a small scale (assuming a Northern hemisphere ecosystem), we created a temperature gradient in the spatial ecosystem with colder temperatures in the 'North' cells and warmer temperatures in the 'Southern' cells. 
```{r temp-viz, fig.cap="Temperature gradient applied to our ecosystem. Red dot indicates the fishing port location."}
temp <- read.csv("./InputData/SardineBanks-Temperature_BaseYear.csv")
temp <- temp %>% pivot_longer(cols=X1:X20)
temp <- temp %>% 
  mutate(name=gsub(name, pattern="X", replacement="")) %>% 
  mutate(name= as.numeric(name),
         X = as.numeric(X)) %>% 
  rename(col=name,
         row=X)
port = tibble(col=10, row=20, value=0)
temp %>% ggplot(aes(x=col, y=rev(row), fill=value)) +
  geom_tile() +
  ylab("") +
  xlab("") + 
  scale_x_continuous(expand=c(0,0)) + 
  scale_y_continuous(expand=c(0,0)) + 
  theme_classic() + 
  geom_point(data=port, aes(x=col, y=rev(row)), size=4, color="red") + 
  labs(fill="Temperature", size="Port") + 
  NULL
```

To ensure the effect we are focusing on is not driven by pre-existing traits related to habitat for foraging, we remove the habitat drivers from functional groups foraging behaviour. The result is that the species are constrained and influenced by environmental (i.e., temperature and primary productivity) drivers rather than limited by habitat type. This included removing the coastal margin of the ecospace to not limit functional groups from the edge of the area. 


```{r depth-viz, eval=F}
depth <- read.csv("./InputData/SardineBanks-Depth.csv")
depth <- depth %>% pivot_longer(cols=X1:X20)
depth <- depth %>% 
  mutate(name=gsub(name, pattern="X", replacement="")) %>% 
  mutate(name= as.numeric(name),
         X = as.numeric(X)) %>% 
  rename(col=name,
         row=X)
depth %>% ggplot(aes(x=col, y=rev(row), fill=value)) +
  geom_tile() +
  ylab("") +
  xlab("") + 
  labs(fill="deptherature") + 
  scale_x_continuous(expand=c(0,0)) + 
  scale_y_continuous(expand=c(0,0)) + 
  theme_classic() + 
  NULL
```


An important part of the ecospace parameterization is the determination of dispersal values. Dispersal determines the average distance a functional group can travel in a year. As it determines how mobile or sessile a functional group is, it is expected to heavily influence the results of marine protected area simulations. The values for similar functional groups were used from [@beattie_marine_2001], and anchovy, benthos, and zooplankton were assumed to have the same values as herring, 'other macrobenthos, and copepods, respectively (Table \@ref(tab:disp)). The dispersal values of all functional groups were modified for the sensitivity analysis. 


```{r disp}
disp <- read.csv("./InputData/Anchovy Bay-Dispersal.csv")
disp <- disp %>% select(Group.name, Base.dispersal.rate)
colnames(disp) = c("Functional Group", "Base dispersal rate (km/year)")
kable(disp, caption="Dispersal values used for Ecospace parameterization.")
```

The fishing fleets had a cost function applied based on distance, and their fishing effort is determined internally through a profit maximizing mechanism. Therefore, they exploit areas of high abundance over areas of low biomass given equal distances. For our specification, we had one port at the Northern end of the habitat halfway through the area (Figure \@ref(fig:temp-viz)). 

We used this ecospace scenario with no MPA as the baseline scenario for all simulations. Then, we applied the described scenarios with the alternative climate change effects and MPA designs. 

We conducted student t-tests to compare biomass, catch, and revenue under RCP 8.5 between various MPA scenarios and the baseline no-MPA scenario. We present significant levels in the figures in the results section and supplemental figures.

### Limitations 
Our work uses a simple ecosystem model with a climate change driver. The model allows us to explore tradeoffs between temperature, fishery pressures, and ecosystems in marine protected area design. This research is not an empirical analysis, thus the results should be interpreted as demonstrative of potential principles and outcomes rather than a strictly literal interpretation. 

Our model includes the spillover of species (i.e. if fish biomass increased in the MPA due to lack of fishing pressure, there can be a positive net migration of fish from no-take areas to the surrounding areas outside the MPA)  based on the dispersal values used (and tested in our sensitivity analysis). The spillover in our model is demonstrated to be nontrivial based on dispersal values used. In the sensitivity analysis (see supplemental results), we vary the dispersal parameter to test varying degrees of spillover, and our results remain robust. 


## Supplemental results


```{r ecosystem-wide-results-from-paper, eval=T}

baseline_decade <- average_data %>% 
  group_by(RCP, design, year, type) %>% 
  summarize(value_aggregate = sum(value)) %>% 
  filter(year <=9) %>% 
  group_by(RCP, design, type) %>% 
  summarize(average_2000s = mean(value_aggregate),
            stdev_2000s = sd(value_aggregate)) %>%
  mutate(se_2000s = stdev_2000s/sqrt(10))

end_of_century <- average_data %>% 
  group_by(RCP, design, year, type) %>% 
  summarize(value_aggregate = sum(value)) %>% 
  filter(year >=90 & year <99) %>% 
  group_by(RCP, design, type) %>% 
  summarize(average_2090s = mean(value_aggregate),
            stdev_2090s = sd(value_aggregate)) %>%
  mutate(se_2090s = stdev_2090s/sqrt(9))


summary_table_revenue1 <- average_fleet_data %>% 
  group_by(RCP, design, year, type) %>% 
  summarize(value_aggregate = sum(value)) %>% 
  filter(year <= 9) %>% 
  group_by(RCP, design, type) %>% 
  summarize(average_2090s = mean(value_aggregate),
            stdev_2090s = sd(value_aggregate)) %>%
  mutate(se_2090s = stdev_2090s/sqrt(10))

summary_table_revenue2 <- average_fleet_data %>% 
  group_by(RCP, design, year, type) %>% 
  summarize(value_aggregate = sum(value)) %>% 
  filter(year >=90 & year <99) %>% 
  group_by(RCP, design, type) %>% 
  summarize(average_2090s = mean(value_aggregate),
            stdev_2090s = sd(value_aggregate)) %>%
  mutate(se_2090s = stdev_2090s/sqrt(9))



summary_table_temp_baseline <- baseline_decade %>% gather(key= variable, value=value, -RCP, -type, -design)
summary_table_temp_end_of_century <- end_of_century %>% gather(key= variable, value=value, -RCP, -type, -design)
summary_table_temp <- rbind(summary_table_temp_baseline, summary_table_temp_end_of_century)
summary_table_temp$type <- gsub(summary_table_temp$type, pattern=".*_", replacement="")
summary_table_temp$year <- gsub(summary_table_temp$variable, pattern=".*_", replacement="")
summary_table_temp$variable <- gsub(summary_table_temp$variable, pattern="_.*", replacement="")
summary_table_temp$aggregate <- paste(summary_table_temp$RCP, summary_table_temp$design, summary_table_temp$year, sep="/")
summary_table_temp$label <- paste(summary_table_temp$design, summary_table_temp$year, sep=" ")

summary_table <- summary_table_temp %>%
  filter(variable != "se")  %>%
  spread(key=variable, value=value)

summary_table_biomass <- summary_table %>% filter(type=="Biomass")
summary_table_catch <- summary_table %>% filter(type=="Catch")
summary_table_effort <- summary_table %>% filter(type=="Effort")

output_table_biomass <- summary_table_biomass %>%
  filter(year=="2090s") %>%
  mutate(value = paste(round(average, 1), " (", round(stdev,2), ")", sep="")) %>%
  select(-c( year,aggregate, label, average, stdev)) %>%
  spread(design, value=value) 

output_table_catch <- summary_table_catch %>%
  filter(year=="2090s") %>%
  mutate(value = paste(round(average, 1), " (", round(stdev,2), ")", sep="")) %>%
  select(-c( year,aggregate, label, average, stdev)) %>%
  spread(design, value=value) 

output_table_summary <- rbind(output_table_biomass, output_table_catch)

output_table_summary <- output_table_summary %>% ungroup()
colnames(output_table_summary) <- c("Climate scenario", "Measure", "Horizontal static", "Network shifting", "Network static", "Square shifting", "Square static", "Vertical static", "No MPA")

kable(output_table_summary, caption="Aggregate results at the end of the century (2090-2099) for all MPA and climate scenarios. Standard deviations included in brackets. All units are in t/km^2^.")


```


```{r bar-chart-catch, fig.cap="Percent change in catch by fleet and target functional groups compared to No MPA scenario at the end of the 21^st^ century. Error bars represent 95% confidence intervals of the mean percent change."}

excluded_fgs <- c("Zooplankton", "Benthos", "Phytoplankton", "Detritus")

decade1 <- seq(0,9,1)
decade2 <- seq(45,54,1)
decade3 <- seq(90,99,1)

average_data <- average_data %>% 
  mutate(TimePeriod = if_else(year %in% decade1, "Today", 
                              if_else(year %in% decade2, "Mid-Century", 
                                      if_else(year %in% decade3, "End of Century", ""))))

average_data$TimePeriod <- fct_relevel(average_data$TimePeriod, "Today", "Mid-Century", "End of Century")

average_data_summary <- average_data %>%
  filter(TimePeriod != "") %>% 
  filter(!group %in% excluded_fgs) %>% 
  filter(RCP=="RCP8.5") %>% 
  group_by(group, TimePeriod, type, design, RCP) %>%
  summarize(stdev = sd(value),
            value=mean(value))

nompa_average <- average_data_summary %>% 
  ungroup() %>% 
  filter(design=="NoMPA")  %>% 
  rename(baseline=value,
         baseline_stdev = stdev) %>% 
  select(-design)
sd_num <- 1.96
average_data_summary <- average_data_summary %>% 
  ungroup() %>% 
  filter(design!="NoMPA") %>% 
  filter(TimePeriod != "Today") %>% 
  filter(TimePeriod != "Mid-Century") %>% 
  left_join(nompa_average) %>% 
  mutate(upper_bound = value + (sd_num*stdev),
         lower_bound = value - (sd_num*stdev)) %>% 
  mutate(perc_change = ((value-baseline)/baseline)*100,
         upper_bound = ((upper_bound-baseline)/baseline)*100,
         lower_bound = ((lower_bound-baseline)/baseline)*100
         )


average_data_summary$group <- gsub(average_data_summary$group, pattern=" ad", replacement= " adult")

average_data_summary$group <- fct_relevel(average_data_summary$group, c("Seiners | Mackerel adult", "Foragers | Anchovy",  "Shrimpers | Shrimp", "Trawlers | Whiting",  "Trawlers | Cod", "Sealers | Seals"), after=0)
catch_bar_plot <- average_data_summary %>% 
  left_join(mpa_labels) %>% 
  filter(type=="Average_Catch") %>% 
  ggplot(aes(x=mpa_label, y=perc_change, fill=perc_change)) +
  geom_col() +
  geom_errorbar(aes(x=mpa_label, ymin=lower_bound, ymax=upper_bound)) + 
  coord_flip() +
  xlab("MPA design") + 
  ylab("Percent change in catch") + 
  facet_wrap(~group, ncol=2, scales="free_x") + 
  scale_fill_gradientn(colors=rev(zissou_continuous)) + 
  labs(fill="Percent change") + 
  NULL

catch_bar_plot
ggsave("./VisualOutputs/Change in catch by scenario under RCP8.5_AlternativeFormat.png", dpi=300, width=8.5, height=11, units="in")

```


At the aggregate level, the fishing effort expended in the ecosystem experiences almost no change from the baseline scenario to the MPA scenarios (Figure \@ref(fig:cpue-plot)). Effort outside MPA experiences a higher percentage increase than that in adjacent cells, which suggests the density of fish may be higher in adjacent cells, thus lowering effort. However, the effort is spatially redirected based on the cost of fishing (i.e., distance from port) and 'drawn' to areas with high biomass. This is apparent when viewing the results of the network MPAs with large edges that can be exploited by fishers. This likely makes these less effective at protecting their biomass, but improves the catch per unit effort in these areas compared to the baseline scenario. 

```{r cpue-plot, fig.cap="Percent change in catch per unit effort (CPUE) by region compared to No MPA scenario at the end of the 21st century."}

fleet_fg_table <- average_fleet_data %>% 
  select(fg, Fleet) %>% 
  unique()

mpa_spatial <- spatial_data %>% 
  left_join(mpa_data) %>% 
  left_join(mpa_legend)

effort_spat <- mpa_spatial %>% 
  filter(type=="Effort") %>% 
  rename(Effort=value,
         Fleet=fg) %>% 
  left_join(fleet_fg_table) %>% 
  select(-type)
catch_spat <- mpa_spatial %>% 
  filter(type=="Catch") %>% 
  rename(Catch=value) %>% 
  select(-type)
  
  
cpue <- catch_spat %>% 
  left_join(effort_spat) %>% 
  mutate(CPUE = Catch/Effort)

cpue_baseline <- cpue %>% 
  filter(design=="NoMPA") %>% 
  select(-c(design, Catch, Effort, mpa, MPA)) %>% 
  rename(Baseline=CPUE)

cpue <- cpue %>% 
  filter(design!="NoMPA") %>% 
  left_join(cpue_baseline) %>% 
  mutate(PercChange= 100*((CPUE-Baseline)/Baseline))
#PercChange produces lots of NaN results because there is no value of effort or catch inside MPAs
cpue %>% 
  group_by(design, MPA, RCP) %>% 
  summarize(PercChange=mean(PercChange, na.rm=TRUE)) %>% 
  ungroup() %>% 
  filter(MPA!="Inside") %>% 
  left_join(mpa_labels) %>% 
  ggplot(aes(x=MPA, y=PercChange, fill=mpa_label)) +
  geom_col(position="dodge") + 
  labs(x = "Region", y = "Percent change (%)", fill = "MPA design") + 
  facet_wrap(~RCP) + 
  NULL

ggsave("./VisualOutputs/CPUE by RCP and MPA Design.png", dpi=300, width=8.5, height=11, units="in")


```


### Sensitivity analysis
When we alter the dispersal parameter, we find the results at the functional group level are altered only slightly at the end of the century (Figure \@ref(fig:sa-dispersal-fg)). In addition, when we look at the area with the most change, we see little variation in catch or biomass levels at the end of the century in cells adjacent to MPAs which theoretically would be most affected by this dispersal parameter (Figure \@ref(fig:sa-spillover)).  


```{r sa-dispersal-fg, echo=F, message=F, warning=F, fig.cap="Biomass by functional group (total for the ecosystem) under alternative spillover values compared to No MPA scenario at the end of the 21st century.", fig.width=11, fig.height=8}

sa_disp <- data.table::fread("/nfs/mpasandclimatechange-data/Data/AnchovyBay/DataOutputs/SA_SpatialData_tidy.csv")
sa_disp <- sa_disp %>% rename(group=fg)
sa_disp <- sa_disp %>% left_join(mpa_data)

excluded_fgs <- c("Zooplankton", "Benthos", "Phytoplankton", "Detritus")

decade1 <- seq(0,9,1)
decade2 <- seq(45,54,1)
decade3 <- seq(90,99,1)

sa_disp <- sa_disp %>% 
  mutate(TimePeriod = if_else(year %in% decade1, "Today", 
                              if_else(year %in% decade2, "Mid-Century", 
                                      if_else(year %in% decade3, "End of Century", ""))))

sa_disp$TimePeriod <- fct_relevel(sa_disp$TimePeriod, "Today", "Mid-Century", "End of Century")

sa_disp$dispersal <- fct_relevel(sa_disp$dispersal, "neg50", "neg25", "0", "pos25", "pos50")
sa_disp$dispersal <- fct_relabel(sa_disp$dispersal, gsub, pattern="pos", replacement="Positive ")

sa_disp$dispersal <- fct_relabel(sa_disp$dispersal, gsub, pattern="neg", replacement="Negative ")
#unique(sa_disp$dispersal)

sa_disp$group <- gsub(sa_disp$group, pattern=" ad", replacement= " adult")
sa_disp$group <- gsub(sa_disp$group, pattern=" juv", replacement= " juvenile")

sa_summary <- sa_disp %>%
  filter(TimePeriod != "") %>% 
  #filter(design %in% design_chosen) %>% 
  filter(!group %in% excluded_fgs) %>% 
  filter(RCP=="RCP8.5") %>% 
  group_by(group, TimePeriod, type, design, RCP, dispersal) %>%
  summarize(value=mean(value))

nompa_average <- sa_summary %>% 
  ungroup() %>% 
  filter(design=="NoMPA")  %>% 
  rename(baseline=value) %>% 
  select(-design)

sa_summary <- sa_summary %>% ungroup() %>% 
  filter(design!="NoMPA") %>% 
  filter(TimePeriod != "Today") %>% 
  filter(TimePeriod != "Mid-Century") %>% 
  left_join(nompa_average) %>% 
  mutate(value= ((value-baseline)/baseline)*100)

biomass_sa_plot <- sa_summary %>% 
  left_join(mpa_labels) %>% 
  filter(type=="Biomass") %>% 
  ggplot(aes(x=mpa_label, y=value, fill=value)) +
  geom_col() +
  coord_flip() +
  xlab("MPA design") + 
  ylab("Percent change in biomass") + 
  facet_wrap(~group +dispersal, ncol=5) + 
  scale_fill_gradientn(colors=rev(zissou_continuous)) + 
  geom_hline(yintercept = 0, lty='dashed', color = 'grey50') + 
  labs(fill="Percent change") + 
  NULL
biomass_sa_plot
ggsave("./VisualOutputs/SA_biomass_dispersal.png", dpi=1200)

```


```{r sa-spillover, fig.cap="Catch and biomass in cells adjacent to MPAs under alternative spillover values at the end of the 21^st^ century.", fig.width=9, fig.height=6, dpi=600}
spillover_effect <- sa_disp %>% 
  filter(year >90) %>% 
  filter(mpa==2) %>% 
  filter(type != "Effort") %>% 
  group_by(type, design, dispersal, mpa) %>% 
  summarize(value=mean(value))

spillover_effect %>% 
  left_join(mpa_labels) %>% 
  ggplot(aes(x=dispersal, y=value, fill=type)) +
  geom_col() +
  #geom_errorbar(aes(x=dispersal, y=value, ymin=value-st_dev, ymax= value+st_dev)) +
  facet_wrap(~mpa_label + type, scales="free_y") +
  ylab(expression("Tonnes / km" ^2)) +
  xlab("Dispersal Change") + 
  theme(legend.position = "None") + 
  NULL
ggsave("./VisualOutputs/SA_spillover_effect.png", dpi=1200)
```


## Supplemental figures

```{r mpa-annotate}

#Create manual borders for each MPA design
horiz_static <- tibble(design="MPA_Horiz_Static", mx1=3, mx2=16, my1=3, my2=5, ax1=1, ax2=2, ay1=1, ay2=2)
network_static1 <- tibble(design="MPA_Network_Static", mx1=8, mx2=13, my1=3, my2=4, ax1=0, ax2=2, ay1=0, ay2=2)
network_static2 <- tibble(design="MPA_Network_Static", mx1=8, mx2=13, my1=6, my2=7, ax1=0, ax2=2, ay1=0, ay2=2)
network_static3 <- tibble(design="MPA_Network_Static", mx1=8, mx2=13, my1=9, my2=11, ax1=0, ax2=2, ay1=0, ay2=2)
network_shifting1 <- tibble(design="MPA_Network_Shifting", mx1=8, mx2=13, my1=2, my2=3, ax1=0, ax2=2, ay1=0, ay2=2)
network_shifting2 <- tibble(design="MPA_Network_Shifting", mx1=8, mx2=13, my1=5, my2=6, ax1=0, ax2=2, ay1=0, ay2=2)
network_shifting3 <- tibble(design="MPA_Network_Shifting", mx1=8, mx2=13, my1=8, my2=10, ax1=0, ax2=2, ay1=0, ay2=2)
square_static <- tibble(design="MPA_Square_Static", mx1=8, mx2=13, my1=3, my2=9, ax1=0, ax2=2, ay1=0, ay2=2)
square_shifting <- tibble(design="MPA_Square_Shifting", mx1=8, mx2=13, my1=2, my2=8, ax1=0, ax2=2, ay1=0, ay2=2)
vert_static <- tibble(design="MPA_Vert_Static", mx1=9, mx2=11, my1=2, my2=15, ax1=0, ax2=2, ay1=0, ay2=2)
mpa_annotation_df <- bind_rows(horiz_static, network_static1, network_shifting1, network_static2, network_shifting2, network_static3, network_shifting3, square_static, square_shifting, vert_static)

#Then need to expand to touch the 'edges' of the MPA rather than start at the center of the cell. 
#Subtract 0.5 from min coords and add 0.5 to max coords. 
mpa_annotation_df$mx1 <- mpa_annotation_df$mx1 - 0.5
mpa_annotation_df$mx2 <- mpa_annotation_df$mx2 + 0.5
mpa_annotation_df$my1 <- mpa_annotation_df$my1 - 0.5
mpa_annotation_df$my2 <- mpa_annotation_df$my2 + 0.5

#I reverse the orientation on the map so I need to reverse the y coordinates here too:
mpa_annotation_df$my1 <- 21  - mpa_annotation_df$my1
mpa_annotation_df$my2 <- 21  - mpa_annotation_df$my2

mpa_annotation_df$ax1 <- mpa_annotation_df$mx1 - 1
mpa_annotation_df$ax2 <- mpa_annotation_df$mx2 + 1
mpa_annotation_df$ay1 <- mpa_annotation_df$my1 + 1
mpa_annotation_df$ay2 <- mpa_annotation_df$my2 - 1

#Make sure the border goes around all of the MPA (important for the network MPA designs)
mpa_annotation_df <- mpa_annotation_df %>% 
  group_by(design) %>% 
  mutate(ax1 = min(ax1),
         ay1 = max(ay1),
         ax2 = max(ax2),
         ay2 = min(ay2)) %>% 
  ungroup() 


nompa <- tibble(design="NoMPA", mx1=NA, mx2=NA, my1=NA, my2=NA, ax1=NA, ax2=NA, ay1=NA, ay2=NA)
mpa_annotation_df <- bind_rows(mpa_annotation_df, nompa) %>% 
  left_join(mpa_labels, by= "design") %>% 
  select(-design) %>% 
  rename(design=mpa_label)


```


```{r whiting-biomass-catch-spatial-plot, eval=F, fig.cap="Biomass and catch of Whiting at the end of the century (Years 90-100)", dpi=600, fig.width=20, fig.height=6}

mpa_spatial <- left_join(spatial_data, mpa_data) 
mpa_spatial$mpa[which(mpa_spatial$design=="NoMPA")] <- 0
mpa_spatial <- mpa_spatial %>% 
  left_join(mpa_legend)
mpa_spatial$MPA <- as.factor(mpa_spatial$MPA)
mpa_spatial$MPA <- fct_relevel(mpa_spatial$MPA, "Outside", after=0)
mpa_spatial$MPA <- fct_relevel(mpa_spatial$MPA, "Inside", after=1)
mpa_spatial <- mpa_spatial %>% 
  left_join(mpa_labels) %>% 
  select(-design) %>% 
  rename(design=mpa_label)

catch_fgs <- c("Whiting")
RCP_val <- "RCP8.5"
mpa_spatial_biomass_fisheries <- end_of_century_average(df=mpa_spatial, groups=catch_fgs)
mpa_spatial_catch_fisheries <- end_of_century_average(df=mpa_spatial, groups=catch_fgs)


max_value <- max(mpa_spatial_biomass_fisheries %>% filter(type=="Biomass") %>% pull(value))
min_value <- min(mpa_spatial_biomass_fisheries %>% filter(type=="Biomass") %>% pull(value)) 
min_value <- floor(min_value)
if (ceiling(max_value)!=1){
  max_value <- ceiling(max_value)
}


mpa_spatial_biomass_fisheries$design <- fct_relevel(mpa_spatial_biomass_fisheries$design, "No MPA", after=0)
mpa_designs <- unique(mpa_spatial_biomass_fisheries$design)
for (design in mpa_designs){
  p <- spatial_maps_segment_plots(df=mpa_spatial_biomass_fisheries, MPA_design=design, type_measure="Biomass", RCP_scenario=RCP_val)
  plot_name <- paste0(design, "Biomass - Whiting")
  assign(plot_name, p)
}

p <- p +
  theme(legend.position = "right") +
  labs(fill=expression("Biomass t/km"^2))
biomass_segment_legend <- cowplot::get_legend(p)

combined_biomass <- cowplot::plot_grid(`No MPABiomass - Whiting`, `Horizontal StaticBiomass - Whiting`, `Vertical StaticBiomass - Whiting`, `Square StaticBiomass - Whiting`, `Square ShiftingBiomass - Whiting`, `Network StaticBiomass - Whiting`, `Network ShiftingBiomass - Whiting`, biomass_segment_legend,  align="h", nrow=1)
#print(combined_biomass)
ggsave("./VisualOutputs/ChangeinBiomassFisheriesGroupsCombined.png", plot=combined_biomass, dpi=1200, height = 4, width=20)


max_value <- max(mpa_spatial_catch_fisheries %>% filter(type=="Catch") %>% pull(value))
min_value <- min(mpa_spatial_catch_fisheries %>% filter(type=="Catch") %>% pull(value)) 
min_value <- floor(min_value)
if (ceiling(max_value)!=1){
  max_value <- ceiling(max_value)
}
for (design in mpa_designs){
  p <- spatial_maps_segment_plots(df=mpa_spatial_catch_fisheries, MPA_design=design, type_measure="Catch", RCP_scenario=RCP_val)
  plot_name <- paste0(design, "Catch - Whiting")
  assign(plot_name, p)
}

p <- p +
  theme(legend.position = "right") +
  labs(fill=expression("Catch t/km"^2))
catch_segment_legend <- cowplot::get_legend(p)


combined_catch <- cowplot::plot_grid(`No MPACatch - Whiting`, `Horizontal StaticCatch - Whiting`, `Vertical StaticCatch - Whiting`, `Square StaticCatch - Whiting`, `Square ShiftingCatch - Whiting`, `Network StaticCatch - Whiting`, `Network ShiftingCatch - Whiting`, catch_segment_legend,
                                   align="h", nrow=1)
#print(combined_catch)
ggsave("./VisualOutputs/ChangeinCatchFisheriesGroupsCombined.png", plot=combined_catch, dpi=1200, height = 4, width=20)


combined_both <- cowplot::plot_grid(`No MPABiomass - Whiting`, `Horizontal StaticBiomass - Whiting`, `Vertical StaticBiomass - Whiting`, `Square StaticBiomass - Whiting`, `Square ShiftingBiomass - Whiting`, `Network StaticBiomass - Whiting`, `Network ShiftingBiomass - Whiting`, biomass_segment_legend,
                                   `No MPACatch - Whiting`, `Horizontal StaticCatch - Whiting`, `Vertical StaticCatch - Whiting`, `Square StaticCatch - Whiting`, `Square ShiftingCatch - Whiting`, `Network StaticCatch - Whiting`, `Network ShiftingCatch - Whiting`, catch_segment_legend,
                                   align="h", nrow=2)
print(combined_both)
ggsave("./VisualOutputs/ChangeinBiomassCatchFisheriesGroupsCombined.png", dpi=1200, height = 8, width=20)


```



```{r functional-groups-biomass, fig.cap="Biomass at the end of the 21^st^ century for each functional group.", fig.width= 20, fig.height=8, dpi=600}
mpa_spatial <- left_join(spatial_data, mpa_data) 
mpa_spatial$mpa[which(mpa_spatial$design=="NoMPA")] <- 0
mpa_spatial <- mpa_spatial %>% 
  left_join(mpa_legend)
mpa_spatial$MPA <- as.factor(mpa_spatial$MPA)
mpa_spatial$MPA <- fct_relevel(mpa_spatial$MPA, "Outside", after=0)
mpa_spatial$MPA <- fct_relevel(mpa_spatial$MPA, "Inside", after=1)
mpa_spatial <- mpa_spatial %>% 
  left_join(mpa_labels) %>% 
  select(-design) %>% 
  rename(design=mpa_label)
mpa_spatial$design <- fct_relevel(mpa_spatial$design, "No MPA", after=0)
mpa_spatial <- mpa_spatial %>% arrange(design)

fgs <- unique(mpa_spatial$fg)
fgs <- fgs[!grepl(fgs, pattern="ers")] #Eliminate fleets so we're just talking about functional groups. 
RCPs <- unique(mpa_spatial$RCP)
mpa_designs <- unique(mpa_spatial$design)

for (fg_val in fgs){
  plot_names <- list()
  spatial_biomass_fg_data <- end_of_century_average(df=mpa_spatial %>% filter(type=="Biomass"), groups = fg_val)
  
  max_value <- max(spatial_biomass_fg_data %>% pull(value))
  min_value <- min(spatial_biomass_fg_data %>% pull(value))
  min_value <- floor(min_value)
  if (ceiling(max_value)!=1){
    max_value <- ceiling(max_value)
  }
  
  for (RCP_val in RCPs){
    for (design in mpa_designs){
      p <- spatial_maps_segment_plots(df=spatial_biomass_fg_data, 
                                      MPA_design=design, 
                                      type_measure="Biomass", 
                                      RCP_scenario = RCP_val)
      if(design == "No MPA" & RCP_val == "RCP2.6"){
        p <- p +
          ggtitle(label=fg_val) +
          NULL
      }
      plot_name <- paste0(fg_val, design, RCP_val, "Biomass")
      plot_names <- append(plot_names, plot_name)
      assign(plot_name, p)
    }
  }
  legend_plot <- p +
    theme(legend.position = "right") +
    labs(fill=expression("Biomass t/km"^2))
  biomass_fg_segment_legend <- cowplot::get_legend(legend_plot)
  plot_list <- mget(unlist(plot_names))
  fg_biomass_plot <- cowplot::plot_grid(plotlist = plot_list, align="h", nrow=2)
  fg_biomass_plot <- cowplot::plot_grid(fg_biomass_plot, biomass_fg_segment_legend, rel_widths = c(10,1),  align="h", nrow=1)
  print(fg_biomass_plot)
  ggsave(paste0("./VisualOutputs/FGBiomassPlot", fg_val, ".png"), dpi=1200, height = 8, width=20)
}

```


```{r functional-groups-catch, fig.cap="Catch at the end of the 21^st^ century for each functional group.", fig.width= 20, fig.height=8, dpi=600}

for (fg_val in catch_fgs){
  plot_names <- list()
  spatial_catch_fg_data <- end_of_century_average(df=mpa_spatial %>% filter(type=="Catch"), groups = fg_val)
  max_value <- max(spatial_catch_fg_data %>% pull(value))
  min_value <- min(spatial_catch_fg_data %>% pull(value))
  min_value <- floor(min_value)
  if (ceiling(max_value)!=1){
    max_value <- ceiling(max_value)
  }
  for (RCP_val in RCPs){
    for (design in mpa_designs){
      p <- spatial_maps_segment_plots(df=spatial_catch_fg_data, 
                                      MPA_design=design, 
                                      type_measure="Catch", 
                                      RCP_scenario = RCP_val)
      if(design == "No MPA" & RCP_val == "RCP2.6"){
        p <- p +
          ggtitle(label=fg_val) +
          NULL
      }
      plot_name <- paste0(fg_val, design, RCP_val, "Catch")
      plot_names <- append(plot_names, plot_name)
      assign(plot_name, p)
    }
  }
  legend_plot <- p +
    theme(legend.position = "right") +
    labs(fill=expression("Catch t/km"^2))
  catch_fg_segment_legend <- cowplot::get_legend(legend_plot)
  plot_list <- mget(unlist(plot_names))
  fg_catch_plot <- cowplot::plot_grid(plotlist = plot_list, align="h", nrow=2)
  fg_catch_plot <- cowplot::plot_grid(fg_catch_plot, catch_fg_segment_legend, rel_widths = c(10,1),  align="h", nrow=1)
  print(fg_catch_plot)
  ggsave(paste0("./VisualOutputs/FGCatchPlot", fg_val, ".png"), dpi=1200, height = 8, width=20)
}

```

